### 2023-07-10 paper.tex

\documentclass[pageno]{jpaper}

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{XXX}

\usepackage[normalem]{ulem}

\begin{document}

\title{
Static Race Detection in OS Kernels by Mining Locking Rules}

\date{}
\maketitle

\thispagestyle{empty}

\begin{abstract}

To take advantage of multiple-core architecture of modern CPUs, an operating 
system (OS) is often designed to be highly concurrent, and processes shared 
data structure in parallel. To ensure safe access to shared data structures, 
fine-grained locking mechanisms are used to avoid data races among different 
threads. However, it is hard to determine whether a lock should be used for a 
specific data structure field (locking rules) even for an expert developer, due 
to poor documentation and complicated logic of OS code. As a result, OS kernel 
is prone to data races, because necessary locks may be missed by mistake. 
Static analysis is a common technique to help improve code quality, but it is 
quite challenging to detect data races in OS kernels automatically, because of 
lack of knowledge of locking rules and high complexity of concurrent execution.

In this paper, we design a practical static analysis approach named RaceMiner, 
to effectively detect harmful races that can trigger memory or logical bugs in 
OS kernels by mining locking rules. RaceMiner first employs an alias-aware rule 
mining method to automatically deduce locking rules, and detects data races 
caused by violation of these rules. And then performs a lock-usage analysis to 
filter out false positives caused by concurrency. At last, RaceMiner extracts 
harmful data races from all detected data races through pattern-based 
estimation. We have evaluated RaceMiner on Linux 6.2, and find xxx data races, 
with a false positive rate of xxx. Among these data races, xxx are estimated to 
be harmful. We have reported these harmful bugs to Linux kernel developers, and 
xxx of them have been confirmed.

\end{abstract}

\input{1.introduction}

\bibliographystyle{plain}
\bibliography{references}

\end{document}