### 2023-07-14 2.motivation.tex

% !TeX spellcheck = en_US
\section{Motivation}
\label{sec_motivation}