### 2023-07-14 4.framework.tex

% !TeX spellcheck = en_US
\section{Framework}
\label{sec_framework}