### 2023-07-27 4.framework.tex
Finish the draft of technique

% !TeX spellcheck = en_US
\section{\sys Approach}
\label{sec_framework}