### 2023-07-27 5.evaluation.tex

% !TeX spellcheck = en_US
\section{Evaluation}
\label{sec_evaluation}

To validate the effectiveness of \sys, we evaluate it on the code of Linux 
kernel 6.2. We run the evaluation on a regular x86-64 desktop with sixteen 
Intel i7-10700 CPU@2.90GHz processors and 64GB physical memory. We use the 
kernel configuration {\em allyesconfig} to enable all kernel code for the 
x86-64 architecture.

\begin{table}[tbph]
	\tablecaption{Detection results of Linux 6.2.}
	\label{tbl_bug_detection}
	\renewcommand{\arraystretch}{1}
	\setlength\tabcolsep{2pt}
	\noindent{\scriptsize
		\begin{center}
			\begin{tabular}{p{1.5cm}|l|c}
				\hline
				\multicolumn{2}{c|}{\textbf{Description}} & \textbf{\sys}  
				\\ \hline
				\multirow{2}{1.5cm}{\textbf{{\em Code analysis}}} & 
				Source files (analyzed/all) & xxxK/xxxK
				\\ \cline{2-3}
				& Source code lines (analyzed/all) & xxxM/xxxM 
				\\ \cline{2-3}
				\hline
				\multirow{3}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
				Key fields / total variables& 
				\\ \cline{2-3}
				& Mined locking rules &
				\\ \cline{2-3}
				& Protected field accesses / all accesses & 0.6
				\\ \cline{2-3}
				\hline
				\multirow{2}{1.5cm}{\textbf{{\em Data race detection}}} & 
				Detected data races (real / all) & 257 / 341
				\\ \cline{2-3}
				& Dropped data races by lock-usage analysis & 59
				\\ \cline{2-3}
				\hline
				\multirow{5}{1.5cm}{\textbf{{\em Data race estimation}}}
				& Null-pointer dereference (confirmed / all) & 15 / 20
				\\ \cline{2-3}
				& Infinite loop (confirmed / all) & 0 / 1
				\\ \cline{2-3}
				& Data inconsistency (confirmed / all) & 7 / 10
				\\ \cline{2-3}
				& Unprotected write (confirmed / all) & 10 / 57
				\\ \cline{2-3}
				& Total harmful data races (confirmed / all) & 32 / 88
				\\ \cline{2-3}
				\hline
				\multirow{3}{1.5cm}{\textbf{{\em Time usage}}} & 
				Key-field extraction & 
				\\ \cline{2-3}
				& Data-race detection &
				\\ \cline{2-3}
				& Total time &
				\\ \cline{2-3}
				\hline
			\end{tabular}
	\end{center}}
\end{table}

\subsection{Bug Detection}
\label{subsec_bug_detection}

We configure \sys with common lock-acquiring/release functions (like {\tt 
spin\_lock and spin\_unlock}) to perform lock-set analysis to extract key 
fields and mine locking rules, and lock-initialization functions (like {\tt 
spin\_lock\_init}) to filter our false data races caused by code paths that 
can not execute concurrently. And then run \sys to automatically check the 
kernel source code. We manually check all the data races found by \sys, and 
Table~\ref{tbl_bug_detection} shows the results, and source code lines are 
counted by CLOC~\cite{cloc}. From the results, we have the following findings:

\PP{Code analysis.} \sys can scale to large code bases of OS kernels, and it in 
total analyze xxxM lines of code in xxxK source code files within xxx hours. 
The remaining xxxM lines of code in xxxK source files are not analyzed, as they 
are not enabled by the {\em allyesconfig} for the x86-64 architecture. We 
believe that \sys can also find more data races in other architectures with 
proper configuration.

\PP{Locking-rule mining.} An OS kernel has a large code base with numerous 
variables. Handling all variables when mining locking rules can introduce much 
overhead. However, we observe that a variable tend to be protected by the lock 
stored in the same data structure as the accessed variable. Based on this 
observation, our locking-rule mining method first extract a key field by 
finding whether there exists any access to it that is protected by a lock 
stored in the same data structure. This method drops xxx\% variables (xxx out 
xxx) that need to handled when mining locking rules, and thus can reduce 
overhead significantly. After extracting key fields, our locking-rule mining 
method collect all accessed to these key fields, and then deducing locking-rule 
based on statistic. In this paper, given a data structure field, we set the 
threshold of the ratio of accesses protected by a specific lock to all accesses 
to 0.6. Our alias-aware rule mining method can drop many false rules and thus 
only mines xxx rules, which can effectively reduce false data races.

\PP{Data race detection.}

\PP{Data race estimation.} Many data races are benign and can not cause memory 
or logic bugs, and thus developers are unwilling to put effort into repairing 
them. We exploit four patterns to detect null-pointer dereferences, infinite 
loop, data inconsistency and unprotected write as introduced in 
Section~\ref{subsec_estimation}, and find 88 data races in these patterns. We 
report them to developers and 32 of them have been confirmed and fixed by them. 
We still wait for response of other data races. Moreover, one of the bluetooth 
l2cap protocol wonder if \sys can be used in their CI to detect these problems. 
The results show that our pattern-based estimation can extract harmful data 
races effectively, and can considerably reduce the workload of developers.