### 2023-07-28 5.evaluation.tex
udpate

% !TeX spellcheck = en_US
\section{Evaluation}
\label{sec_evaluation}

To validate the effectiveness of \sys, we evaluate it on the code of Linux 
kernel 6.2. We run the evaluation on a regular x86-64 desktop with sixteen 
Intel i7-10700 CPU@2.90GHz processors and 64GB physical memory. We use the 
kernel configuration {\em allyesconfig} to enable all kernel code for the 
x86-64 architecture.

\begin{table}[tbph]
	\tablecaption{Detection results of Linux 6.2.}
	\label{tbl_bug_detection}
	\renewcommand{\arraystretch}{1}
	\setlength\tabcolsep{2pt}
	\noindent{\scriptsize
		\begin{center}
			\begin{tabular}{p{1.5cm}|l|c}
				\hline
				\multicolumn{2}{c|}{\textbf{Description}} & \textbf{\sys}  
				\\ \hline
				\multirow{2}{1.5cm}{\textbf{{\em Code analysis}}} & 
				Source files (analyzed/all) & xxxK/xxxK
				\\ \cline{2-3}
				& Source code lines (analyzed/all) & xxxM/xxxM 
				\\ \cline{2-3}
				\hline
				\multirow{3}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
				Key fields / total variables& 
				\\ \cline{2-3}
				& Mined locking rules &
				\\ \cline{2-3}
				& Protected field accesses / all accesses & 0.6
				\\ \cline{2-3}
				\hline
				\multirow{2}{1.5cm}{\textbf{{\em Data race detection}}} & 
				Detected data races (real / all) & 273 / 341
				\\ \cline{2-3}
				& Dropped data races by lock-usage analysis & 63
				\\ \cline{2-3}
				\hline
				\multirow{5}{1.5cm}{\textbf{{\em Data race estimation}}}
				& Null-pointer dereference (confirmed / all) & 15 / 20
				\\ \cline{2-3}
				& Infinite loop (confirmed / all) & 0 / 1
				\\ \cline{2-3}
				& Data inconsistency (confirmed / all) & 7 / 10
				\\ \cline{2-3}
				& Unprotected write (confirmed / all) & 10 / 57
				\\ \cline{2-3}
				& Total harmful data races (confirmed / all) & 32 / 88
				\\ \cline{2-3}
				\hline
				\multirow{3}{1.5cm}{\textbf{{\em Time usage}}} & 
				Key-field extraction & 
				\\ \cline{2-3}
				& Data-race detection &
				\\ \cline{2-3}
				& Total time &
				\\ \cline{2-3}
				\hline
			\end{tabular}
	\end{center}}
\end{table}

\subsection{Bug Detection}
\label{subsec_bug_detection}

We configure \sys with common lock-acquiring/release functions (like {\tt 
spin\_lock and spin\_unlock}) to perform lock-set analysis to extract key 
fields and mine locking rules, and lock-initialization functions (like {\tt 
spin\_lock\_init}) to filter our false data races caused by code paths that 
can not execute concurrently. And then run \sys to automatically check the 
kernel source code. We manually check all the data races found by \sys, and 
Table~\ref{tbl_bug_detection} shows the results, and source code lines are 
counted by CLOC~\cite{cloc}. From the results, we have the following findings:

\PP{Code analysis.} \sys can scale to large code bases of OS kernels, and it in 
total analyze xxxM lines of code in xxxK source code files within xxx hours. 
The remaining xxxM lines of code in xxxK source files are not analyzed, as they 
are not enabled by the {\em allyesconfig} for the x86-64 architecture. We 
believe that \sys can also find more data races in other architectures with 
proper configuration.

\PP{Locking-rule mining.} An OS kernel has a large code base with numerous 
variables. Handling all variables when mining locking rules can introduce much 
overhead. However, we observe that a variable tend to be protected by the lock 
stored in the same data structure as the accessed variable. Based on this 
observation, our locking-rule mining method first extract a key field by 
finding whether there exists any access to it that is protected by a lock 
stored in the same data structure. This method drops xxx\% variables (xxx out 
xxx) that need to handled when mining locking rules, and thus can reduce 
overhead significantly. After extracting key fields, our locking-rule mining 
method collect all accessed to these key fields, and then deducing locking-rule 
based on statistic. In this paper, given a data structure field, we set the 
threshold of the ratio of accesses protected by a specific lock to all accesses 
to 0.6. Our alias-aware rule mining method can drop many false rules and thus 
only mines xxx rules, which can effectively reduce false data races.

\PP{Data race detection.} \sys reports 341 data races in the kernel source 
code. We spent 15 hours on checking these data races and identify that 273 of 
them are real, with a false positive rate of 19.9\%. Besides, our lock-usage 
analysis drops 63 false data races and the results shows that it can reduce 
false positives significantly.

\PP{Data race estimation.} Many data races are benign and can not cause memory 
or logic bugs, and thus developers are unwilling to put effort into repairing 
them. We exploit four patterns to detect null-pointer dereferences, infinite 
loop, data inconsistency and unprotected write as introduced in 
Section~\ref{subsec_estimation}, and find 88 data races in these patterns. We 
report them to developers and 32 of them have been confirmed and fixed by them. 
We still wait for response of other data races. Moreover, one of the developers 
wonders if \sys can be used in their CI to detect these problems. 
The results show that our pattern-based estimation can extract harmful data 
races effectively, and can considerably reduce the workload of developers.

\subsection{False Positives and Negatives}
\label{subsec_false_pos_neg}
\sys reports 68 false data races, and through manually checking these false 
data races, we find that they are introduced for three main reasons:

First, \sys employs an alias-aware rule mining method to deduce locking rules. 
To improve the precision of mined rules, it assumes that the accessed variable 
and the protecting lock are stored in the same data structure. Even so, \sys 
can also deduce some false rules because some developers does not use locks 
properly. They take a lock/unlock pair to protect accesses to all fields in the 
same data structure, instead of the exact field should be protected for 
convenience. As a result, our alias-aware rule mining method infers that 
accesses to all fields surrounded by the lock/unlock pair should be protected 
by the lock. This reason causes \sys to report 47 false data races.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1\linewidth]{figures/fig_demo_false_rule.pdf}
	\figcaption{A false data race caused by an incorrect locking rule.}
	\label{fig_demo_false_rule}
\end{figure}

Figure~\ref{fig_demo_false_rule} shows a false data race caused by an incorrect 
locking rule. In this example, only accesses to {\em cmd->t\_state} and {\em 
cmd->transport\_state} should be protected by the lock {\em 
cmd->t\_state\_lock}. However, the lock operation is put ahead of the switch 
statement by developers, making our alias-aware rule mining method deduces that 
{\em cmd->scsi\_status} and {\em cmd->se\_cmd\_flags} also need to be protected 
by {\em cmd->t\_state\_lock} mistakenly. Based on this incorrect locking rule, 
\sys reports a false positive at Line 745 when {\em cmd->se\_cmd\_flags} is 
accessed in an if statement. Although this data race is a false positive, it 
can cause performance degradation because the critical zone should have been 
limited to Lines 899-900.

Second, in order to reduce memory overhead and improve the performance of data 
passing among different functions, an integer can be divided into several bit 
vector to represent different data structure fields. However,