### 2023-07-31 8.conclusion.tex

% !TeX spellcheck = en_US
\section{Conclusion}
\label{sec_conclusion}

In this paper, we develop a novel static approach named \sys to detect data 
races in OS kernels by mining locking rules. It consists of three key 
techniques, including an alias-aware rule mining method to automatically deduce 
locking rules, a lock-usage analysis to filter out false positives caused by 
code paths that can not execute concurrently and a pattern-based estimation to 
extract harmful data races that can trigger memory or logical bugs such as 
null-pointer dereferences and data inconsistency. In the evaluation, \sys finds 
88 real harmful data races, and 32 of them have been confirmed by the 
developers.