### 2023-08-02 2.motivation.tex

% !TeX spellcheck = en_US
\section{Motivation}
\label{sec_motivation}

\subsection{A Motivating Example}
\label{subsec_motivating_example}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1\linewidth]{figures/fig_demo_bug.pdf}
	\figcaption{A null-pointer dereference due to data race in Linux 6.2.}
	\label{fig_demo_bug}
\end{figure}

Figure~\ref{fig_demo_bug} shows a real null-pointer dereference caused by a 
data race in the Linux DRM driver. In this example, We exploit 
\textcircled{\footnotesize{n}} to represent the execution order of instructions 
and show one execution case on the left of Figure~\ref{fig_demo_bug}. In the 
DRM driver, the functions {\tt exynos\_drm\_crtc\_atomic\_disable()} and {\tt 
exynos\_crtc\_handle\_event()} can execute concurrently. In Thread 1, the 
variables {\tt crtc->state->event} and {\tt crtc->state->active} are checked by 
an if statement in the function {\tt exynos\_drm\_crtc\_atomic\_disable()}. 
After the condition is calculated to be true, the function {\tt 
exynos\_crtc\_handle\_event()} execute in Thread 2. In this function, the value 
of {\tt crtc->state->event} is assigned to {\tt event} 
(\textcircled{\footnotesize{2}}) and then {\tt event} is checked in an if 
statement (\textcircled{\footnotesize{3}}). If it is not NULL, the variable 
{\tt crtc->state->event} is assigned with NULL 
(\textcircled{\footnotesize{4}}). Right after this assignment, the function 
{\tt drm\_crtc\_send\_vblank\_event()} is called  in Thread 1 
(\textcircled{\footnotesize{6}}) with the argument {\tt crtc->state->event}, 
after acquiring the lock {\tt crtc->dev->event\_lock} 
(\textcircled{\footnotesize{5}}). In the called function, the variable {\tt 
crtc->state->event} is finally dereferenced through {\tt e->pipe} 
(\textcircled{\footnotesize{7}}). In this execution case, the variable {\tt 
crtc->state->event} is first assigned with a NULL value and then dereferenced, 
and thus a null-pointer dereference can occur.

This bug is triggered only when {\tt crtc->state->event} is set to NULL by 
Thread 2 right after the first condition of the if statement in Thread 1 is 
calculated to be true. Such a requirement is difficult to satisfy by executing 
existing test suites. In fact, this bug had existed for nearly 6 years since 
Linux 4.14 (Released in Nov. 2017), and it was fixed by us based on a report 
generated by \sys.

\subsection{Challenges}
\label{subsec_challenges}
Detecting data races and estimating their harmfulness in OS kernels have three 
main challenges:

\PP{C1: Getting locking rules.} The relationship between variables and locks is 
not well documented in OS kernels, making it hard to determine whether a 
specific variable should be protected by a lock and which lock is required, 
even for an expert developer. And thus existing 
annotation-basedapproaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, 
Anderson:PLDI09, Zhou:MICRO19, Flanagan:PASTE01, Flanagan:PLDI00, 
Sadowski:PLATEAU14, ClangThreadSafety, Blackshear:OOPSLA18} are difficult to 
apply to race detection in OS kernels. Other approaches~\cite{Choi:PLDI02, 
Engler:SOSP03, Voung:FSE07, Pratikakis:PLDI06, Naik:PLDI06} employ 
lockset-based analysis to detect data races automatically, but they do not 
consider alias relationships~\cite{Voung:FSE07, Engler:SOSP03} or just use 
imprecise flow-insensitive alias analysis~\cite{Choi:PLDI02, 	
Pratikakis:PLDI06, Naik:PLDI06}. However, due to the heavy use of pointers and 
data structure fields in kernel code, the alias relationships between variables 
can be very complex, and thus lacking effective alias analysis can introduce 
both false positives and false negatives.

\PP{C2: Dropping false data races.} Due to lack of run-time information, static 
analysis suffers from false positives. For example, each data race involves 
more than one code path that should be able to concurrently execute. However, 
which code can execute concurrently is not well documented for an OS kernel, 
and it is also hard to determine concurrent code statically due to the 
complexity of OS code. Thus static analysis can report many false data races.

\PP{C3: Estimating harmfulness of data races.} Many data races are benign or 
introduced by developers deliberately to improve the performance of OS kernels, 
and can not cause memory or logic bugs in fact. Therefore, developers are 
unwilling to put effort into repairing these data races. To automatically 
detect harmful data races, most approaches~\cite{Narayanasamy:PLDI07, 
Sen:PLDI08, Kasikci:SOSP13, Kasikci:ASPLOS12} explore thread inter-leavings 
dynamically to trigger data races and observe their impact on the checked 
software. However, they suffer from low code coverage and thus can miss many 
real harmful data races.