### 2023-08-02 paper.tex

\documentclass[pageno]{jpaper}

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{XXX}

\usepackage[normalem]{ulem}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{xspace}
\usepackage{xstring}

\usepackage[numbers, sort]{natbib}

\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

\newcommand{\sys}{\mbox{RaceMiner}\xspace}

\newcommand{\figcaption}{% 
	\setlength{\abovecaptionskip}{4pt}%	\setlength{\belowcaptionskip}{1pt}% 
	\caption}

\newcommand{\tablecaption}{% 
	\setlength{\abovecaptionskip}{6pt}%
	\setlength{\belowcaptionskip}{2pt}% 
	\caption}

\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}

\newcommand{\PP}[1]{
	\vspace{2px}
	\noindent{\bf \IfEndWith{#1}{.}{#1}{#1.}}
}

\begin{document}

\title{
Static Race Detection in OS Kernels by Mining Locking Rules}

\date{}
\maketitle

\thispagestyle{empty}

\begin{abstract}

To take advantage of multiple-core architecture of modern CPUs, an operating 
system (OS) is expected to handle concurrency, inevitably introducing many 
concurrency issues. And data race is one of the most harmful concurrency issues 
that can trigger memory or logic bugs. Data races are caused by wrongly 
accessing shared data in different threads. To ensure correct access to shared 
data, fine-grained locking mechanisms are introduced. However, it is hard to 
determine whether a lock should be held when accessing a specific variable 
(locking rules) even for an expert developer, because of poor document and 
complicated logic of OS code. As a result, OS kernel is prone to data races, 
because necessary locks may be missed by mistake. Static analysis is a common 
technique to improve code quality, but it is quite challenging to detect data 
races in OS kernels automatically, due to lack of knowledge of locking rules 
and high complexity of concurrent execution.

In this paper, we design a practical static analysis approach named \sys, to 
effectively detect data races and estimate their harmfulness in OS kernels by 
mining locking rules. \sys first employs an alias-aware rule mining method to 
automatically deduce locking rules, and detects data races caused by violation 
of these rules. And then performs a lock-usage analysis to filter out false 
positives caused by code that can not execute concurrently. At last, \sys 
extracts harmful data races from all detected data races through a 
pattern-based estimation. We have evaluated \sys on Linux 6.2, and find 273 
data races, with a false positive rate of 19.9\%. Among these data races, 88 
are estimated to be harmful. We have reported these harmful data races to Linux 
kernel developers, and 32 of them have been confirmed.

\end{abstract}

\input{1.introduction}
\input{2.motivation}
\input{3.technique}
\input{4.framework}
\input{5.evaluation}
\input{6.discussion}
\input{7.related}
\input{8.conclusion}

\footnotesize
\bibliographystyle{plain}
\bibliography{references}

\end{document}